Perturbing SU($n$)-symmetric interactions

### 2020-01-28 collective_sun/sun_perturbation.tex
fixed minor error in multi-body eigenvalue problem
simplifications to derivations of M-body eigenvalue problem and the two-body product

To compute the effective Hamiltonian induced by the single-body
perturbation $\V_1$, we use the permutational symmetry of the
ground-state manifold to simplify
\begin{align}
  H_1^{(1)} = \sum_{X,p} v_{Xp} \P_0 X_p \P_0
  \EQFS \sum_{X,p} v_{Xp} X_0
  \EQFS \sum_X \bar v_X \col{X},
  \label{eq:H_1_1}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, and
\begin{align}
  \bar v_X \equiv \f1N \sum_p v_{Xp},
  &&
  \col{X} \equiv \sum_p X_p,
\end{align}
are respectively the mean value in $\v v_X$ and the collective version
of $X$.  In order to compute the second order effective Hamiltonian
$H_1^{(2)}$, we first choose an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V1_psi}):
\begin{align}
  H_0 \V_1 \ket\psi
  = E_0 \V_1 \ket\psi
  + \sum_X \p{\sum_{p,q} h_{pq} v_{Xp} X_q - \sum_p h_p v_{Xp} X_p}
  \ket\psi,
  &&
  h_p \equiv \sum_q h_{pq},
  \label{eq:H_V1_psi}
\end{align}
where we define $h_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We thus find that a
perturbation $\V_1^\Delta\ket\psi$ (indexed by $\Delta$) is an
eigenvector of the interaction Hamiltonian $H_0$ if its coefficient
vectors $\v v_X^\Delta$ all satisfy the eigenvalue equation
\begin{align}
  \p{\m h - \diag\v h}\c\v v = \Delta \v v,
  &&
  \m h \equiv \sum_{p,q} h_{pq} \op{p}{q},
  &&
  \v h \equiv \sum_q h_q \ket{q},
  \label{eq:cond_1}
\end{align}
where $\m h$ is a matrix of all $h_{pq}$, and
$\diag\v h\equiv\sum_p h_p \op{p}$ is a matrix with $\v h$ on the
diagonal and zeros everywhere else.  Solving the general eigenvalue
problem in \eqref{eq:cond_1} requires diagonalizing the $N\times N$
matrix $\m h-\diag\v h$.  When the interaction Hamiltonian $H_0$ is
translationally invariant, however, this eigenvalue problem can be
solved analytically; we provide the corresponding solution in Appendix
\ref{sec:trans_inv}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Diagnosing perturbed states with the interaction Hamiltonian}

Here we simplify vectors of the form $H_0\V_M\ket\psi$ to arrive at
the expansions in \eqref{eq:H_V1_psi} and \eqref{eq:H_V2_psi}.

@@

We wish to simplify
\begin{align}
  H_0 \V_1 \ket\psi
  = \sum_k \sum_{p<q} h_{pq} v_k \Pi_{pq} X_k \ket\psi,
  \label{eq:H_V1_psi_start}
\end{align}
where $\ket\psi\in\M_\FS$ is an arbitrary fully symmetric state, and
without loss of generality we consider a single-body perturbation
$\V_1$ built out of only one single-body operator $X$.  The sum in
\eqref{eq:H_V1_psi_start} has terms with $k\in\set{p,q}$, and terms
with $k\notin\set{p,q}$.  In the case of $k\notin\set{p,q}$, the
permutation operator $\Pi_{pq}$ commutes with $X_k$ and annihilates on
$\ket\psi$, leaving terms at each fixed $k$ of the form
\begin{align}
  \sum_{\substack{p<q\\k\notin\set{p,q}}} h_{pq}
  = \sum_{p<q} h_{pq} - \sum_{\substack{p<q\\k\in\set{p,q}}} h_{pq}
  = E_0 - \f12 \sum_{\substack{p,q\\k\in\set{p,q}}} h_{pq}
  = E_0 - h_k,
\end{align}
where we define $h_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $h_{pq}=h_{qp}$ and $h_{pp}=0$; and
$h_k \equiv \sum_\ell h_{k\ell}$.  The terms in
\eqref{eq:H_V1_psi_start} with $k\notin\set{p,q}$ are then
\begin{align}
  \sum_{\substack{p<q\\k\notin\set{p,q}}}
  h_{pq} v_k \Pi_{pq} X_k \ket\psi
  = \sum_k \p{E_0 - h_k} v_k X_k \ket\psi
  = E_0 \V_1 \ket\psi - \sum_k h_k v_k X_k \ket\psi
\end{align}
while the terms in \eqref{eq:H_V1_psi_start} with $k\in\set{p,q}$ take
the form
\begin{align}
  \sum_{\substack{p<q\\k\in\set{p,q}}}
  h_{pq} v_k \Pi_{pq} X_k \ket\psi
  = \sum_{k<q} h_{kq} v_k X_q \ket\psi
  + \sum_{p<k} h_{pk} v_k X_p \ket\psi
  = \sum_{k,q} h_{kq} v_k X_q \ket\psi
\end{align}
which implies
\begin{align}
  H_0 \V_1 \ket\psi
  = E_0 \V_1 \ket\psi
  + \p{\sum_{p,q} h_{pq} v_p X_q - \sum_p h_p v_p X_p}
  \ket\psi.
\end{align}

@@

Given an $M$-body perturbation
\begin{align}
  \V_M \equiv \sum_{k\in\C_N\p{M}} w_k O_k,
\end{align}
where $O$ is an $M$-body operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
a choice of $M$ distinct spins that $O$ acts on; and $\C_N\p{n}$ is a
choice of $n$ elements from a set of $N$; we wish to simplify
\begin{align}
  H_0 \V_M \ket\psi
  = \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\in\C_N\p{M}}} h_{pq} w_k
  \Pi_{pq} O_k \ket\psi,
  \label{eq:H_VM_psi_start}
\end{align}
where $\ket\psi\in\M_\FS$ is an arbitrary fully symmetric state.  The
sum in \eqref{eq:H_VM_psi_start} has terms with $p,q\in k$, terms with
$p,q\notin k$, and terms with only one of $p$ or $q\in k$.  The
permutation operator $\Pi_{pq}$ acts trivially on $O_k\ket\psi$ when
$p,q\in k$ or $p,q\notin k$, leaving terms at each fixed $k$ of the
form
\begin{align}
  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\in k~\t{or}~p,q\notin k}} h_{pq}
  = \sum_{\p{p,q}\in\C_N\p{2}} h_{pq}
  - \sum_{p\in k} \sum_{\substack{q\in\ZZ_N\\q\notin k}} h_{pq}
  = E_0 - \sum_{p\in k}
  \p{\sum_{q\in\ZZ_N} h_{pq} - \sum_{q\in k} h_{pq}},
\end{align}
where we first split a sum over $\p{p,q}\in\C_N\p{2}$ with the
restriction that $p,q\in k~\t{or}~p,q\notin k$ into an unrestricted
sum minus a remainder, and then similarly split the sum over all spins
$q$ with the restriction that $q\notin k$ into a sum over all $q$
minus the remainder $q\in k$.  Defining $h_p\equiv\sum_q h_{pq}$, we thus find that
\begin{align}
  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\in k~\t{or}~p,q\notin k}} h_{pq}
  = E_0 + 2\sum_{\p{p,q}\subset k} h_{pq} - \sum_{p\in k} h_p,
\end{align}
where the sum over $\p{p,q}\subset k$ denotes a sum over distinct
choices of two elements in $k$.  The terms in
\eqref{eq:H_VM_psi_start} with only one of $p\in k$ or $q\in k$,
meanwhile, are
\begin{align}
  \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{q\in k}
  h_{pq} w_k \Pi_{pq} O_k \ket\psi
  = \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
  \sum_{j=1}^M h_{p k_j} w_k \Pi_{p k_j} O_k \ket\psi.
\end{align}
Defining
$k_{jp}\equiv\p{k_1,k_2,\cdots,k_{j-1},p,k_{j+1},\cdots,k_M}$, i.e.~an
element of $\C_N\p{M}$ in which the $j$-th number $k_j$ has been
replaced by $p$, we can simplify
\begin{align}
  \Pi_{p k_j} O_k \ket\psi = O_{k_{jp}}\ket\psi,
\end{align}
and switch the labels for $p$ and $k_j$ to find that
\begin{align}
  \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{q\in k}
  h_{pq} w_k  \Pi_{pq} O_k \ket\psi
  = \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
  \sum_{j=1}^M h_{pk_j} w_{k_{jp}} O_k \ket\psi.
\end{align}
Defining
\begin{align}
  \p{h\circ w}_k
  \equiv \sum_{j=1}^M \sum_{p\in k} h_{pk_j} w_{k_{jp}}
  = h_{pk_1} w_{pk_2k_3\cdots k_M}
  + h_{pk_2} w_{k_1pk_3\cdots k_M}
  + h_{pk_M} \cdots w_{k_1k_2\cdots k_{M-1} p}
\end{align}
and
\begin{align}
  \bar h_k \equiv 2 \sum_{\p{p,q}\subset k} h_{pq} - \sum_{p\in k} h_p,
\end{align}
we thus find that
\begin{align}
  H_0 \V_M \ket\psi
  = E_0 \V_M \ket\psi + \sum_{k\in\C_N\p{M}}
  \sp{\bar h_k w_k + \p{h\circ w}_k} O_k \ket\psi.
\end{align}

@@

Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
\eqref{eq:perturbations} requires us to simplify
\begin{align}
  \P_0 \p{\V_2^\Delta}^2 \P_0
  = \sum_{O,Q} \sum_{\substack{k<\ell\\p<q}}
  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
  \EQFS \f14 \sum_{O,Q} \sum_{k,\ell,p,q}
  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq},
  \label{eq:POQP_start}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$; we define $w_{Opq}$ for all $p,q$ (as
opposed to only $p<q$) by enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$;
and the coefficient vectors $\v w_O$ satisfy the eigenvalue equation
\eqref{eq:cond_2} with eigenvalue $\Delta\ne0$, which implies that all
$\v{\m w}_O$ are mean-zero.  To simplify the sum in
\eqref{eq:POQP_start}, we first consider the terms with
$\p{p,q}=\p{k,\ell}$:
\begin{align}
  \sum_{k<\ell}
  w_{Ok\ell} w_{Qk\ell} O_{k\ell} Q_{k\ell}
  = \v{\m w}_O \c \v{\m w}_Q O_{0,1} Q_{0,1}.
\end{align}
We then consider terms $p\in\set{k,\ell}$ and $q\notin\set{k,\ell}$ or
vice versa:
\begin{align}
  \f14 \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
  &\EQFS \f14 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qkq} \p{O_{0,1} Q_{0,2} + O_{0,1} Q_{1,2}} \\
  \f14 \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
  &\EQFS \f14 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qkq} \p{O_{0,1} Q_{2,0} + O_{0,1} Q_{2,1}},
\end{align}
where
\begin{align}
  \f14 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}} w_{Ok\ell} w_{Qkq}
  = \f14 \sum_{k,\ell,q} w_{Ok\ell} w_{Qkq}
  - \f14 \sum_{k,\ell} w_{Ok\ell} w_{Qk\ell}
  = \f14 \v w_O\c\v w_Q - \f12 \v{\m w}_O \c \v{\m w}_Q.
\end{align}
Finally, the terms in \eqref{eq:POQP_start} with
$p,q\notin\set{k,\ell}$ are
\begin{align}
  \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
  \EQFS \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq} O_{0,1} Q_{2,3},
\end{align}
where, using the fact that all $\v{\m w}_O$ are mean-zero,
\begin{multline}
  \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}} w_{Ok\ell} w_{Qpq}
  = - \f14 \sum_{\substack{k,\ell\\p,q\in\set{k,\ell}}} w_{Ok\ell} w_{Qpq}
  - \f14 \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq}
  - \f14 \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq} \\
  = -\v w_O \c \v w_Q + \v{\m w}_O \c \v{\m w}_Q.
\end{multline}
Altogether, we thus have that
\begin{multline}
  \P_0 \p{\V_2^\Delta}^2 \P_0
  \EQFS - \sum_{O,Q} \bigg[
  \p{\v w_O \c \v w_Q - \v{\m w}_O \c \v{\m w}_Q} O_{0,1} Q_{2,3}
  - \v{\m w}_O \c \v{\m w}_Q O_{0,1} Q_{0,1} \\
  -\f14 \p{\v w_O\c\v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q}
  O_{0,1} \p{Q_{0,2} + Q_{2,0} + Q_{1,2} + Q_{2,1}} \bigg].
  \label{eq:POQP}
\end{multline}
In principle, our next task is to write this result in terms of
collective operators within the fully symmetric manifold; this task
turns out to be rather non-trivial in general.  We therefore discuss
the decomposition of multi-local operators (e.g.~$O_{0,1} Q_{1,2}$, a
3-local operator) into a sum of collective operators within the fully
symmetric manifold in Appendix \ref{sec:multi_to_collective}.  At the
end of Appendix \ref{sec:multi_to_collective}, we provide analytical
expressions for the special cases of
\begin{enumerate*}
\item multi-local operators that are tensor powers of a single
  operator, i.e.~of the form $X^{\otimes k}$; and
\item multi-local operators of the form $X^{\otimes k}$ with $X^2=1$.
\end{enumerate*}
These cases are relevant for e.g.~the SU(2) case with Ising-type
interactions of the form $Z\otimes Z$, with $Z$ a single-spin Pauli
operator.

>>>>

To compute the effective Hamiltonian induced by the single-body
perturbation $\V_1$, we use the permutational symmetry of the
ground-state manifold to simplify
\begin{align}
  H_1^{(1)} = \sum_{X,p} v_{Xp} \P_0 X_p \P_0
  \EQFS \sum_{X,p} v_{Xp} X_0
  \EQFS \sum_X \bar v_X \col{X},
  \label{eq:H_1_1}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, and
\begin{align}
  \bar v_X \equiv \f1N \sum_p v_{Xp},
  &&
  \col{X} \equiv \sum_p X_p,
\end{align}
are respectively the mean value in $\v v_X$ and the collective version
of $X$.  In order to compute the second order effective Hamiltonian
$H_1^{(2)}$, we first choose an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V1_psi}):
\begin{align}
  H_0 \V_1 \ket\psi
  = E_0 \V_1 \ket\psi
  + \sum_X \sum_p \p{\sum_q h_{pq} v_{Xq} - h_p v_{Xp}} X_p
  \ket\psi,
  &&
  h_p \equiv \sum_q h_{pq},
  \label{eq:H_V1_psi}
\end{align}
where we define $h_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We thus find that a
perturbation $\V_1^\Delta\ket\psi$ (indexed by $\Delta$) is an
eigenvector of the interaction Hamiltonian $H_0$ if its coefficient
vectors $\v v_X^\Delta$ all satisfy the eigenvalue equation
\begin{align}
  \p{\m h - \diag\v h}\c\v v = \Delta \v v,
  &&
  \m h \equiv \sum_{p,q} h_{pq} \op{p}{q},
  &&
  \v h \equiv \sum_q h_q \ket{q},
  \label{eq:cond_1}
\end{align}
where $\m h$ is a matrix of all $h_{pq}$, and
$\diag\v h\equiv\sum_p h_p \op{p}$ is a matrix with $\v h$ on the
diagonal and zeros everywhere else.  Solving the general eigenvalue
problem in \eqref{eq:cond_1} requires diagonalizing the $N\times N$
matrix $\m h-\diag\v h$.  When the interaction Hamiltonian $H_0$ is
translationally invariant, however, this eigenvalue problem can be
solved analytically; we provide the corresponding solution in Appendix
\ref{sec:trans_inv}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Diagnosing perturbed states with the interaction Hamiltonian}
\label{sec:H_VM_psi}

In this section, we simplify the products of the form
$H_0\V_M\ket\psi$ to arrive at the expansions in \eqref{eq:H_V1_psi}
and \eqref{eq:H_V2_psi}.  Here $H_0$ is the SU($n$)-symmetric
interaction Hamiltonian in \eqref{eq:ints}, $\ket\psi$ is a fully
symmetric state, and $\V_M$ is an $M$-body perturbation of the form
\begin{align}
  \V_M \equiv \sum_{k\in\C_N\p{M}} w_k O_k.
\end{align}
where $O$ is an $M$-body operator that is invariant under arbitrary
permutation of its tensor factors; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
a choice of $M$ distinct spins that $O$ acts on; and $\C_N\p{n}$ is a
choice of $n$ elements from a set of $N$.

@@

We first simplify
\begin{align}
  H_0 \V_1 \ket\psi
  = \sum_k \sum_{p<q} h_{pq} v_k \Pi_{pq} X_k \ket\psi,
  \label{eq:H_V1_psi_start}
\end{align}
where without loss of generality we consider a single-body
perturbation $\V_1$ built out of only one single-body operator $X$.
The sum in \eqref{eq:H_V1_psi_start} has terms with $k\in\set{p,q}$,
and terms with $k\notin\set{p,q}$.  In the case of $k\notin\set{p,q}$,
the permutation operator $\Pi_{pq}$ commutes with $X_k$ and
annihilates on $\ket\psi$, leaving terms at each fixed $k$ of the form
\begin{align}
  \sum_{\substack{p<q\\k\notin\set{p,q}}} h_{pq}
  = \sum_{p<q} h_{pq} - \sum_{\substack{p<q\\k\in\set{p,q}}} h_{pq}
  = E_0 - \f12 \sum_{\substack{p,q\\k\in\set{p,q}}} h_{pq}
  = E_0 - h_k,
\end{align}
where we define $h_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $h_{pq}=h_{qp}$ and $h_{pp}=0$; and
$h_k \equiv \sum_\ell h_{k\ell}$.  The terms in
\eqref{eq:H_V1_psi_start} with $k\notin\set{p,q}$ are then
\begin{align}
  \sum_{\substack{p<q\\k\notin\set{p,q}}}
  h_{pq} v_k \Pi_{pq} X_k \ket\psi
  = \sum_k \p{E_0 - h_k} v_k X_k \ket\psi
  = E_0 \V_1 \ket\psi - \sum_k h_k v_k X_k \ket\psi
\end{align}
while the terms in \eqref{eq:H_V1_psi_start} with $k\in\set{p,q}$ take
the form
\begin{align}
  \sum_{\substack{p<q\\k\in\set{p,q}}}
  h_{pq} v_k \Pi_{pq} X_k \ket\psi
  = \sum_{k<q} h_{kq} v_k X_q \ket\psi
  + \sum_{p<k} h_{pk} v_k X_p \ket\psi
  = \sum_{k,q} h_{kq} v_k X_q \ket\psi
\end{align}
which implies
\begin{align}
  H_0 \V_1 \ket\psi
  = E_0 \V_1 \ket\psi
  + \sum_p \p{\sum_{q} h_{pq} v_q - h_p v_p}
  X_p \ket\psi.
\end{align}

@@

We wish to simplify
\begin{align}
  H_0 \V_M \ket\psi
  = \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\in\C_N\p{M}}} h_{pq} w_k
  \Pi_{pq} O_k \ket\psi,
  \label{eq:H_VM_psi_start}
\end{align}
where without loss of generality we consider an $M$-body perturbation
$\V_M$ built out of only one $M$-body operator $O$.  The sum in
\eqref{eq:H_VM_psi_start} has terms with $p,q\in k$, terms with
$p,q\notin k$, and terms with only one of $p$ or $q\in k$.  The
permutation operator $\Pi_{pq}$ acts trivially on $O_k\ket\psi$ when
$p,q\in k$ or $p,q\notin k$, leaving terms at each fixed $k$ of the
form
\begin{align}
  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\in k~\t{or}~p,q\notin k}} h_{pq}
  = \sum_{\p{p,q}\in\C_N\p{2}} h_{pq}
  - \sum_{p\in k} \sum_{\substack{q\in\ZZ_N\\q\notin k}} h_{pq}
  = E_0 - \sum_{p\in k}
  \p{\sum_{q\in\ZZ_N} h_{pq} - \sum_{q\in k} h_{pq}},
\end{align}
where we first split a sum over $\p{p,q}\in\C_N\p{2}$ with the
restriction that $p,q\in k~\t{or}~p,q\notin k$ into an unrestricted
sum minus a remainder, and then similarly split the sum over all spins
$q$ with the restriction that $q\notin k$ into a sum over all $q$
minus the remainder $q\in k$.  We thus find that
\begin{align}
  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\in k~\t{or}~p,q\notin k}} h_{pq}
  = E_0 + \bar h_k,
  &&
  \bar h_k \equiv 2\sum_{\p{p,q}\subset k} h_{pq} - \sum_{p\in k} h_p,
  &&
  h_p \equiv \sum_q h_{pq},
\end{align}
where the sum over $\p{p,q}\subset k$ denotes a sum over distinct
choices of two elements in $k$.  The terms in
\eqref{eq:H_VM_psi_start} with only one of $p\in k$ or $q\in k$,
meanwhile, are
\begin{align}
  \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{q\in k}
  h_{pq} w_k \Pi_{pq} O_k \ket\psi
  = \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
  \sum_{j=1}^M h_{p k_j} w_k \Pi_{p k_j} O_k \ket\psi.
\end{align}
Defining
$k_{jp}\equiv\p{k_1,k_2,\cdots,k_{j-1},p,k_{j+1},\cdots,k_M}$, i.e.~an
element of $\C_N\p{M}$ in which the $j$-th number $k_j$ has been
replaced by $p$, we can simplify
\begin{align}
  \Pi_{p k_j} O_k \ket\psi = O_{k_{jp}}\ket\psi,
\end{align}
and switch the labels for $p$ and $k_j$ to find that
\begin{align}
  \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{q\in k}
  h_{pq} w_k  \Pi_{pq} O_k \ket\psi
  = \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
  \sum_{j=1}^M h_{pk_j} w_{k_{jp}} O_k \ket\psi.
\end{align}
Defining
\begin{align}
  \p{h\circ w}_k
  \equiv \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{j=1}^M
  h_{pk_j} w_{k_{jp}}
  = \sum_{\substack{p\in\ZZ_N\\p\notin k}}
  \p{h_{pk_1} w_{pk_2k_3\cdots k_M}
    + h_{pk_2} w_{k_1pk_3\cdots k_M}
    + \cdots + h_{pk_M} w_{k_1k_2\cdots k_{M-1} p}},
\end{align}
we thus find that
\begin{align}
  H_0 \V_M \ket\psi
  = E_0 \V_M \ket\psi + \sum_{k\in\C_N\p{M}}
  \sp{\bar h_k w_k + \p{h\circ w}_k} O_k \ket\psi.
\end{align}
If $M=1$, then $\bar h_k=-h_k$ and
$\p{h\circ w}_k=\sum_\ell h_{k\ell}w_\ell$, so
\begin{align}
  H_0 \V_1 \ket\psi
  = E_0 \V_1 \ket\psi + \sum_{k\in\ZZ_N}
  \p{\sum_\ell h_{k\ell} w_\ell - h_k w_k} O_k \ket\psi.
\end{align}
If $M=2$, then
$\p{h\circ w}_{pq}=\sum_k\p{h_{kp} w_{kq}+h_{kq}w_{pk}}$, so
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi + \sum_{p<q}
  \sp{\bar h_{pq} w_{pq} + \sum_k\p{h_{kp} w_{kq} + h_{kq} w_{pk}}}
  O_k \ket\psi,
\end{align}
where $\bar h_{pq}=2h_{pq}-h_p-h_q$ with $h_p\equiv\sum_qh_{pq}$.

@@

Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
\eqref{eq:perturbations} requires us to simplify
\begin{align}
  \P_0 \p{\V_2^\Delta}^2 \P_0
  = \sum_{O,Q} \sum_{\substack{k<\ell\\p<q}}
  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
  \EQFS \f14 \sum_{O,Q} \sum_{k,\ell,p,q}
  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq},
  \label{eq:POQP_start}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$; we define $w_{Opq}$ for all $p,q$ (as
opposed to only $p<q$) by enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$;
and the coefficient vectors $\v w_O$ satisfy the eigenvalue equation
\eqref{eq:cond_2} with eigenvalue $\Delta\ne0$, which implies that all
$\v{\m w}_O$ are mean-zero.  To simplify the sum in
\eqref{eq:POQP_start}, we first consider the terms with
$\p{p,q}=\p{k,\ell}$:
\begin{align}
  \sum_{k<\ell}
  w_{Ok\ell} w_{Qk\ell} O_{k\ell} Q_{k\ell}
  = \v{\m w}_O \c \v{\m w}_Q O_{0,1} Q_{0,1}.
\end{align}
We then consider terms $p\in\set{k,\ell}$ and $q\notin\set{k,\ell}$ or
vice versa:
\begin{align}
  \f14 \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
  &\EQFS \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qkq} O_{0,1} Q_{1,2} \\
  \f14 \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
  &\EQFS \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qkq} O_{0,1} Q_{1,2},
\end{align}
where
\begin{align}
  \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}} w_{Ok\ell} w_{Qkq}
  = \sum_{k,\ell,q} w_{Ok\ell} w_{Qkq}
  - \sum_{k,\ell} w_{Ok\ell} w_{Qk\ell}
  = \v w_O\c\v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q.
\end{align}
Finally, the terms in \eqref{eq:POQP_start} with
$p,q\notin\set{k,\ell}$ are
\begin{align}
  \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
  \EQFS \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq} O_{0,1} Q_{2,3},
\end{align}
where, using the fact that all $\v{\m w}_O$ are mean-zero,
\begin{multline}
  \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}} w_{Ok\ell} w_{Qpq}
  = - \f14 \sum_{\substack{k,\ell\\p,q\in\set{k,\ell}}} w_{Ok\ell} w_{Qpq}
  - \f14 \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq}
  - \f14 \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq} \\
  = -\v w_O \c \v w_Q + \v{\m w}_O \c \v{\m w}_Q.
\end{multline}
Altogether, we thus have that
\begin{multline}
  \P_0 \p{\V_2^\Delta}^2 \P_0
  \EQFS \sum_{O,Q}
  \bigg[ \v{\m w}_O \c \v{\m w}_QO_{0,1} Q_{0,1}
  + \p{\v w_O\c\v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q}
  O_{0,1} Q_{1,2} \\
    - \p{\v w_O \c \v w_Q - \v{\m w}_O \c \v{\m w}_Q}
    O_{0,1} Q_{2,3} \bigg].
  \label{eq:POQP}
\end{multline}
In principle, our next task is to write this result in terms of
collective operators within the fully symmetric manifold; this task
turns out to be rather non-trivial in general.  We therefore discuss
the decomposition of multi-local operators (e.g.~$O_{0,1} Q_{1,2}$, a
3-local operator) into a sum of collective operators within the fully
symmetric manifold in Appendix \ref{sec:multi_to_collective}.  At the
end of Appendix \ref{sec:multi_to_collective}, we provide analytical
expressions for the special cases of
\begin{enumerate*}
\item multi-local operators that are tensor powers of a single
  operator, i.e.~of the form $X^{\otimes k}$; and
\item multi-local operators of the form $X^{\otimes k}$ with $X^2=1$.
\end{enumerate*}
These cases are relevant for e.g.~the SU(2) case with Ising-type
interactions of the form $Z\otimes Z$, with $Z$ a single-spin Pauli
operator.
